---
output:
  md_document:
    variant: markdown_github
html_preview: false
---

#  meshes: MeSH Enrichment and Semantic analyses

```{r echo=FALSE, results="hide", message=FALSE}
library("txtplot")
library("ypages")
```

`r badge_bioc_release("meshes", "green")`
`r badge_devel("meshes", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/meshes.svg)](https://www.bioconductor.org/packages/devel/bioc/html/meshes.html#since)
`r badge_bioc_download("meshes", "total", "blue")`
`r badge_bioc_download("meshes", "month", "blue")`


[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/GuangchuangYu/meshes/branch/master/graph/badge.svg)](https://codecov.io/gh/GuangchuangYu/meshes)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/meshes/commits/master)
[![GitHub forks](https://img.shields.io/github/forks/GuangchuangYu/meshes.svg)](https://github.com/GuangchuangYu/meshes/network)
[![GitHub stars](https://img.shields.io/github/stars/GuangchuangYu/meshes.svg)](https://github.com/GuangchuangYu/meshes/stargazers)


[![platform](http://www.bioconductor.org/shields/availability/devel/meshes.svg)](https://www.bioconductor.org/packages/devel/bioc/html/meshes.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/meshes.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/meshes/)
[![Linux/Mac Travis Build Status](https://img.shields.io/travis/GuangchuangYu/meshes/master.svg?label=Mac%20OSX%20%26%20Linux)](https://travis-ci.org/GuangchuangYu/meshes)
[![AppVeyor Build Status](https://img.shields.io/appveyor/ci/Guangchuangyu/meshes/master.svg?label=Windows)](https://ci.appveyor.com/project/GuangchuangYu/meshes)




MeSH (Medical Subject Headings) is the NLM controlled vocabulary used
to manually index articles for MEDLINE/PubMed. MeSH terms were
associated by Entrez Gene ID by three methods, gendoo, gene2pubmed and
RBBH. This association is fundamental for enrichment and semantic
analyses. meshes supports enrichment analysis (over-representation and
gene set enrichment analysis) of gene list or whole expression
profile. The semantic comparisons of MeSH terms provide quantitative
ways to compute similarities between genes and gene groups. meshes
implemented five methods proposed by Resnik, Schlicker, Jiang, Lin and
Wang respectively and supports more than 70 species.



[![Twitter](https://img.shields.io/twitter/url/https/github.com/GuangchuangYu/meshes.svg?style=social)](https://twitter.com/intent/tweet?hashtags=meshes&url=https://guangchuangyu.github.io/meshes&screen_name=guangchuangyu)


----------------------------------------------------------------------------------------

For details, please visit our project website, <https://guangchuangyu.github.io/meshes>.

+ [Documentation](https://guangchuangyu.github.io/meshes/documentation/)
+ [Feedback](https://guangchuangyu.github.io/meshes/#feedback)


### Download stats

[![download](http://www.bioconductor.org/shields/downloads/meshes.svg)](https://bioconductor.org/packages/stats/bioc/meshes/)
`r badge_bioc_download("meshes", "total", "blue")`
`r badge_bioc_download("meshes", "month", "blue")`


```{r echo=F, comment=NA, eval=F}
x <- dlstats::bioc_stats("meshes")
x <- x[-nrow(x), ] # last month data is not complete
txtplot(ggtree::Date2decimal(x$end), x$Nb_of_downloads, width=100, height=25)
```






